---
title: "Chapter 4: Classification"
---

**Stock Market Data**

* For each date, we have recorded the percentage returns for each of the five previous trading days, `Lag1` through `Lag5`. 
* We have also recorded `Volume` (the number of shares traded on the previous day, in billions), `Today` (the percentage return on the date in question) and `Direction` (whether the market was Up or Down on this date). 
* Our goal is to predict Direction (a qualitative response) using the other features.

```{r load_libraries}
library(MASS)
library(ISLR2)
library(stargazer)     # Summarizing
library(tidyverse)     # Data processing
library(ggthemes)      # Plotting
library(car)           # VIF
```

```{r}
names(Smarket)
dim(Smarket)
summary(Smarket)
pairs(Smarket)
```

* The correlations between the lag variables and today’s returns are close to zero. In other words, there appears to be little correlation between today’s returns and previous days’ returns. 
* The only substantial correlation is between `Year` and `Volume`. By plotting the data, which is ordered chronologically, we see that Volume is increasing over time. In other words, the average number of shares traded daily increased from 2001 to 2005.

```{r}
# Direction variable is qualitative
cor(Smarket[, -9])
```

```{r}
attach(Smarket)
plot(Volume)
```

**Logistic Regression**

* The smallest p-value here is associated with `Lag1`. 
* The negative coefficient for this predictor suggests that if the market had a positive return yesterday, then it is less likely to go up today. However, at a value of 0.15, the p-value is still relatively large, and so there is no clear evidence of a real association between `Lag1` and `Direction`.

```{r}
glm.fits <- glm(
  Direction ~ Lag1 + Lag2 + Lag3 + Lag4 + Lag5 + Volume, data = Smarket, 
  family = 'binomial'
)
summary(glm.fits)
```

```{r}
coef(glm.fits)
summary(glm.fits)$coef

# P-values
summary(glm.fits)$coef[, 4]
```

**Logistic Regression Predictions**

* The `predict()` function can be used to predict the probability that the market will go up, given values of the predictors. The type = "response" option tells R to output probabilities of the form P(Y = 1|X), as opposed to other information such as the logit.

```{r}
glm.probs <- predict(glm.fits, type = "response")
glm.probs[1:10]
```

```{r}
contrasts(Direction)
```

```{r}
glm.pred <- rep("Down", 1250)
glm.pred[glm.probs > .5] = "Up"
```

**Confusion Matrix**

```{r}
table(glm.pred, Direction)
```

* Accuracy

```{r}
mean(glm.pred == Direction)
```

**Train-Test Split**

```{r}
train <- (Year < 2005)
Smarket.2005 <- Smarket[!train, ]
dim(Smarket.2005)
Direction.2005 <- Direction[!train]
```

```{r}
glm.fits <- glm(
  Direction ~ Lag1 + Lag2 + Lag3 + Lag4 + Lag5 + Volume, data = Smarket, family = binomial, subset = train
)

glm.probs <- predict(glm.fits, Smarket.2005, type = "response")
```

* The results are rather disappointing: the test error rate is 52%, which is worse than random guessing! Of course this result is not all that surprising, given that one would not generally expect to be able to use previous days’ returns to predict future market performance.

```{r}
glm.pred <- rep("Down", 252)
glm.pred[glm.probs > .5] <- "Up"
table(glm.pred, Direction.2005)

mean(glm.pred == Direction.2005)
```

* Perhaps by removing the variables that appear not to be helpful in predicting Direction, we can obtain a more effective model. After all, using predictors that have no relationship with the response tends to cause a deterioration in the test error rate.
* This suggests a possible trading strategy of buying on days when the model predicts an in- creasing market, and avoiding trades on days when a decrease is predicted. Of course one would need to investigate more carefully whether this small improvement was real or just due to random chance.

```{r}
glm.fits <- glm(Direction ~ Lag1 + Lag2, data = Smarket, family = binomial, subset = train)
glm.probs <- predict(glm.fits, Smarket.2005, type = "response")
glm.pred <- rep("Down", 252)
glm.pred[glm.probs > .5] <- "Up"
table(glm.pred, Direction.2005)
mean(glm.pred == Direction.2005)
```

```{r}
predict(glm.fits, newdata =
          data.frame(Lag1 = c(1.2, 1.5), 
                     Lag2 = c(1.1, -0.8)), 
        type = "response"
)
```

**Linear Discriminant Analysis**

```{r}
lda.fit <- lda(Direction ~ Lag1 + Lag2, data = Smarket, subset = train) 
lda.fit
```

